\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-04-22 paper/paper.tex
If accepted, paper edits.

\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}
We present an approach to modeling failure data with heavy censoring, multiple censoring and multiple truncation. We present the generalized limited failure population model (GFLP) of Chan and Meeker which provides a generative model for failure distributions characterized by a "bathtub" hazard function (citation). This model may provide an adequate fit to failures of complex systems by differentiating between early and late types of failures as well as the propensity toward early failure. We discuss the case where the data are grouped, suggesting that we might expand the model, accounting for differences among groups. We show that by modeling the data hierarchically, we can deal with groups of various sizes by borrowing information across groups. We illustrate our approach on a large set of failure data for hard disk drives made available by Backblaze (citation).

\subsection{Background}
Products can frequently fail due to more than one cause.  For example, there are many parts in a washing machine that can break and cause the whole machine to fail.  Decorative tree lighting can fail if one bulb in the chain burns out.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \cite{nelson}.  A common assumption is the times to failure for each risk, $s$, are statistically independent.  Thus, the overall reliability of a unit can be modeled using the product rule across all $s$ risks.  The GFLP model is a special case of the $s$ independent competing risks model with 2 modes of failure: defects and wearout.  Sometimes the cause of the failure is know; but more often the observations do not contain any information about the failure mode \cite{chan}.  The GFLP model was first used to describe the early and late failure modes of CB radios.  No failure information was available for this data, which made fitting the model a challenge using maximum likelihood estimation.  Other authors, for example, Basu et al., used a Bayesian approach to fit the GFLP model to masked failure data from an engineering system \cite{basu}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.

\subsection{Motivation}
The goal of this paper is to model, and compare, the lifetime distribution of the various hard drive brands using parametric lifetime models.  There are numerous features of the data, however, that make standard estimation approaches problematic.  The first issue is many hard drives have been running for quite some time before entering the sample.  Even hard drives starting after 2013 appear in the data after hundreds of hours of operation.  Also, although we have four years of data, hard drives rarely fail so many units are still in service.  Thus, we have both left truncation and right censoring.  The other challenge is each hard drive model varies in number units in operation and total time on test.  Some models have hundreds of units in service; others have as few as two.  Combined with the truncation this makes stable maximum likelihood (ML) estimates difficult to obtain for many drive models. \\

Another issue is many product populations contain a mixture of defective and nondefective units.  The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.  Ignoring this hazard structure, which is known to exist for computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.  A model that accounts for multiple failure modes is the general limited failure population model (GFLP) that combines different failure time distributions; for example, one failure mode for defective units (infant mortality) and a second mode for wearout \cite{chan}.  Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are unknown, which presents challenges for parameter identification in the GFLP model, as discussed in Chan and Meeker (1999).  Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.  With over 60 brand drive models in testing a hierarchical framework is advantageous as it pools information from across brands to get more precise estimates of the individual parameters.  This is especially helpful for the Backblaze data as there are many drive models with as few as one or two failures.

\subsection{Overview}
The structure of the paper is as follows.  Section 1 presents the Backblaze data and the unique modeling challenges it presents, such as multiple failure modes, right censoring, and left truncation.  Section 2 introduces the GFLP model as a mixture of two Weibull distributions.  Using the Backblaze data, in Section 3 we fit a 5 parameter GFLP model that allows for separate Weibull terms for each failure mode and estimates $p$, the proportion of units subject to infant mortality.  In Section 4 we discuss computational and modeling issues that sugguests a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.  In Section 5 we compare hard drive brand models based on our final model.  In Section 6  we examine the posterior distributions of the model parameters to assess goodness of fit for the hierarchical model.  Finally, in Section 7 we review the GFLP model and propose future extensions.

@@

In Figure 2 we plot the Kaplan-Meier adjusted cdf for hard drive model X on Weibull paper.  Each point on the plot corresponds to a hard drive failure.  Censored drives are not plotted.  As mentioned in the introduction, the population of hard drives exhibits two primary failure modes.  One mode is a result of manufacturing defects, which cause early failures, known as infant mortality.  The second mode is non-defective hard drives that eventually fail due to wearout.   Evidence of at least two failure modes is seen in the Kaplan Meier plot with a kink occurring around hour Z.  Therefore, fitting a single Weibull model would not be flexible enough to model the failure distribution.     [INSERT FIGURE, Do we Want to Put Greenwood Bands on Too?]

@@

>>>>

\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}
We present an approach to modeling failure data with heavy censoring, multiple censoring and multiple truncation. We extend the generalized limited failure population model (GFLP) of Chan and Meeker which provides a generative model for failure distributions characterized by a "bathtub" hazard function (citation). This model provides an adequate fit to failures of complex systems by differentiating between early and late types of failures as well as the propensity toward early failure. We discuss the case where the data are grouped, suggesting that we might expand the model, accounting for differences among groups. We show that by modeling the data hierarchically, we can deal with groups of various sample sizes by borrowing information across groups. We illustrate our approach on a large set of failure data for hard disk drives made available by the hard drive backup company, Backblaze (citation).

\subsection{Background}
Products can frequently fail due to more than one reason.  For example, there are many parts in a washing machine that can break causing the whole machine to fail.  Decorative tree lighting can fail if one bulb in the chain burns out.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \cite{nelson}.  A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.  Thus, the overall reliability of a unit can be modeled using the product rule across all $s$ risks.  The GFLP model is a special case of the $s$ independent competing risks model with 2 modes of failure: defects and wearout.  Sometimes the cause of the failure is know; but more often the observations do not contain information about the failure mode.  The GFLP model was first introduced to describe the early and late failure modes of CB radios \cite{chan}.  No failure information was available for this data, which made estimation a challenge using maximum likelihood estimation.  Other authors, for example, Basu et al., used a Bayesian approach to fit the GFLP model to masked failure data from an engineering system \cite{basu}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.

\subsection{Motivation}
The goal of this paper is to model, and compare, the lifetime distribution of the various hard drive brands using the GFLP lifetime model.  There are numerous features of the data, however, that make standard estimation approaches for this parametric model problematic.  The first issue is many hard drives have been running for quite some time before entering the sample.  Even hard drives starting after 2013 appear in the data after hundreds of hours of operation.  When an observation is at risk before entering the study, this is known as left truncation.  Also, although we have four years of data, hard drives rarely fail so many units are still in service.  Thus, we have both left truncation and right censoring.  The other challenge is each hard drive model varies in number units in operation and total time on test.  Some models have hundreds of units in service; others have as few as two.  Combined with the truncation and censoring this makes stable maximum likelihood (ML) estimates difficult to obtain for many drive models. \\

Another issue is product populations often contain a mixture of defective and nondefective units.  The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.  Ignoring this hazard structure, which is known to exist for computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.  The general limited failure population model (GFLP) combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.  Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are unknown, which presents challenges for parameter identification in the GFLP model, as discussed in Chan and Meeker (1999).  Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.  With over 60 brand drive models in testing a hierarchical framework is advantageous as it pools information from across brands to get more precise estimates of the individual parameters.  This is especially helpful for the Backblaze data as there are many drive models with as few as one or two failures.

\subsection{Overview}
The structure of the paper is as follows.  Section 1 summarizes the Backblaze data and the unique modeling challenges it presents.  Section 2 introduces the GFLP model as a mixture of two Weibull distributions.  In Section 3 we fit multiple specifications of the GFLP model to best model the Backblaze data.  We also discuss computational and modeling issues that suggests a final model with a common Weibull distribution for defective units and individual Weibull distributions for the wearout failure mode.  In Section 4 we compare hard drive brand models based on our final GFLP model.  Finally, in Section 5we review the GFLP model and propose future extensions.

@@

In Figure 2 we plot the Kaplan-Meier adjusted cdf for hard drive model X on Weibull paper.  Each point on the plot corresponds to a hard drive failure.  Censored drives are not plotted.  As mentioned in the introduction, the population of hard drives exhibits two primary failure modes.  One mode is a result of manufacturing defects, which cause early failures, known as infant mortality.  The second mode is non-defective hard drives that eventually fail due to wearout.   Evidence of at least two failure modes is seen in the Kaplan Meier plot with a kink occurring around hour Z.  Therefore, fitting a single Weibull model would not be flexible enough to model the failure distribution.

@@

The set of models to compare was chosen based on the data, interpretation of the model, as well as estimation considerations.  For all model specifications the location and scale parameter of the first failure mode was fixed across drive brand models.  Partly this was due to a lack of early failures in the data.  Moreover, though, from a consumer or business standpoint, comparing the lifetime distribution of infant mortality across hard drive brand models is not particularly informative.  The more practical questions are: is this hard drive likely to suffer from infant morality? and how long should it expect to run before it fails from wearout?  Assuming a constant failure distribution for infant mortality allows the first question to be answered since $\pi$'s are comparable across brands, which would not be the case if each drive brand model had a unique Weibull distribution fit to the early failure mode.  Flexibility in the parameters for the 2nd failure mode is necessary as initial plots of the data indicate heterogeneity across brands in the right tails of the failure distribution.  Going from a fixed model for all hard drive brands and gradually increasing the complexity of the model, we end up at Model 4, which has the most parameters.  Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by brand drive model.
